\chapter*{Acknowledgements}

First of all I want to thank my thesis advisor Erik Adli who hired me and gave me the opportunity to join this exciting research project.
I would also like to thank Patric Muggli, my second thesis advisor, who I also got to share office with for a while.
I have learned a lot from you both.
Thanks to the AWAKE Collaboration for letting me be a part of the early stages of a truly interesting and challenging experiment, and see it through to its first year of operation.

The work presented in this thesis is mainly built on computer simulations.
Access to the simulation software OSIRIS was provided by Instituto Superior Técnico (IST) in Lisbon.
Thanks to Luís Silva for letting me visit several times, and to Jorge Vieira and Ligia Diana Amorim for our collaborations.
I enjoyed my time there, and I absolutely loved Lisbon.
The computing power was provided by the University of Oslo and the Norwegian Metacenter for Computational Science, using the \textit{Abel} computing cluster through project code \texttt{nn9303k}.

During my first year I got to spend a month and a half at the Stanford National Accelerator Laboratory (SLAC) in California, looking over the shoulders of the people running the FACET plasma wakefield experiment.
Thanks to Mark Hogan for letting me visit.
I enjoyed my brief time at FACET and spending time with the fantastic people working there.

A big thank you to Helga Holmestad and Kyrre Sjøbæk for letting me live almost four months in total in your guest room before I moved to CERN for the last year and a half and got my own place.
I also want to thank the AWAKE people at CERN -- especially Edda Gschwendtner, Marlene Turner and Alexey Petrenko -- for making me feel welcome in the group.

A special thank you to LGBTQ CERN and UN-GLOBE Geneva for being my friends and sparkly social network.
Anders, Mario, Alex, Laura, Tess, Ana, Xinyi, Noppadol, Pasqual, Nanoo, and all the others ... it wouldn't have been the same without you!

Finally, a big thank you to the High Energy Physics (EPF) section at the University of Oslo for the many fun years I got to spend there, as well as the Building~6/8/9/10 people at CERN where I've been working while finishing the thesis, with an extra mention of Ryan Bodenstein for proof-reading it.

To all of you, mentioned or not: thanks for all the interesting discussions, drinks, nerding, as well as all the laughs.
It's been a blast.
So long, and thanks for all the fish!

\vfill
\null\hfill\textbf{Veronica Berglyd Olsen}
\newline
\null\hfill\textit{Saint-Genis, July 2018}
