\chapter*{Preface}

Plasma wakefield accelerators are complex. and there are many parameters to tweak in order to accelerate a beam of a quality that is useful for applications in particle physics. medicine, or other types of research. As Terry Pratchett once said: \textit{It is well known that a vital ingredient of success is not knowing that what you're attempting can't be done} \cite{pratchett:1987}. Even if a useful particle collider cannot be built based on a plasma wakefield accelerator design, understanding how charged particle beams interact with plasmas may lead to other applications. The strong beam focusing forces produced by plasmas under certain conditions can be taken advantage of in for instance plasma lenses \cite{su:1990}. Computer simulations are a very useful tool to try and understand complex systems where many factors interact. They can be used to find to study ideal cases, or they can be used to replicate experiments in order to better understand what is going on when you cannot measure all the properties of an experiment.

The work presented in this thesis is aimed towards addressing some of the questions surrounding the design of Run 2 of the AWAKE experiment at CERN (see Chapters \ref{Ch:Intro} and \ref{Ch:WFA}). While Run 1 addresses some of the principle properties of a proton driven plasma wakefield accelerator, like the interaction between the plasma and the beam itself, how the wakefields evolve, and how a sample of electrons behave in such a wakefield, Run 2 aims to accelerate a narrow, short electron beam to high energies while retaining a low energy spread and low emittance. The work is published in the following four papers:

\begin{enumerate}[I]
    \item Loading of a Plasma-Wakefield Accelerator Section Driven by a Self-Modulated Proton Bunch, \emph{Proceedings of IPAC 2015} \cite{berglyd_olsen:2015}
    \item Loading of Wakefields in a Plasma Accelerator Section Driven by a Self-Modulated Proton Beam, \emph{Proceedings of NAPAC 2016} \cite{berglyd_olsen:2016}
    \item Data Acquisition and Controls Integration of the AWAKE Experiment at CERN, \emph{Proceedings of IPAC 2017} \cite{berglyd_olsen:2017}
    \item Emittance Preservation of an Electron Beam in a Loaded Quasilinear Plasma Wakefield, \emph{Physical Review Accelerators and Beams} \cite{berglyd_olsen:2018}
\end{enumerate}

The thesis also includes an introductory chapter outlining some of the core concepts involved in plasma wakefield acceleration techniques in Chapter \ref{Ch:Intro}. The AWAKE experiment itself is outlined in more detail in Chapter \ref{Ch:WFA}. The simulations forming the basis for the publications are described in Chapter \ref{Ch:Sim}, where the approximations used are also described. In Chapter \ref{Ch:DAQ} some of the additional work of integrating the AWAKE experiment with the CERN Control System is outlined. A final summary and conclusions are found in Chapter \ref{Ch:SnC}.

\noindent The four publications are included in this thesis in an appendix titled \hyperref[A:Pub]{Publications}. Additional appendices outlining the principles of Particle in Cell codes used in this work, and the analysis code written, are also included.

\section*{Notation}

\begin{table}[hbt]
    \centering
    \caption{Overview of notation used in this thesis.}
    \label{T:Notes}
    \begin{tabular}{p{0.12\linewidth} p{0.78\linewidth}}
        \rowcolor{tblhead}
        \texthh{Notation}           & \texthh{Description} \\
        \hline
        $n_{0}$                     & The average or initial plasma density \\
        $n_{pe}$                    & The density of plasma electrons \\
        $n_{b}$                     & The density of a general particle beam \\
        $n_{eb}$, $n_{pb}$          & The density of an electron or a proton beam in particular \\
        % \hline
        $\sigma_{r}$                & The width of a Gaussian beam when it is assumed to be round \\
        $\sigma_{x}$, $\sigma_{y}$  & The transverse size of a Gaussian beam when it may not be
                                      round, or the value applies to only one plane. \\
        % \hline
        $\alpha$, $\beta$, $\gamma$ & The Twiss parameters, also known as the Courant-Snyder
                                      parameters \cite{courant:1958}. \\
        $\betar$, $\gammar$         & Relativistic factors* \\
        \hline
    \end{tabular}
\end{table}

\paragraph{*}
These are indexed with \emph{r} in this thesis to separate them from the Twiss parameters.

\vfill